---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Lindsay Stall"
date: "2/7/2023"
output: github_document
---

### Load packages and data


```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r}
# install.packages("devtools")
devtools::install_github("rstudio-education/dsbox")
```


```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

```{r}
view(dennys)
```

```{r}
view(laquinta)
```

### Exercise 1

```{r}
dn_ak <- dennys %>%
  filter(state == "AK")
nrow(dn_ak)
```

There are 3 Denny's locations in Alaska

### Exercise 2

```{r}
lq_ak <- laquinta %>%
  filter(state == "AK")
nrow(lq_ak)
```

There are 2 La Quinta locations in Alaska


### Exercise 3

```{r}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

### Exercise 4 
There are 6 observations and the variables are address, city, zip, longitude, and latitude (of both x and y seperately) and state.


### Exercise 5

```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```


### Exercise 6

```{r}
dn_lq_ak <- dn_lq_ak %>%
  mutate(dn_lq_ak, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_ak
```


### Exercise 7

```{r}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_ak_mindist
```


### Exercise 8

```{r}
dn_lq_ak%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```

```{r}
dn_lq_ak_mindist %>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```


### Exercise 9

```{r}
dn_nc <- dennys %>%
  filter(state == "NC")
nrow(dn_nc)
```

```{r}
lq_nc <- laquinta %>%
  filter(state == "NC")
nrow(lq_nc)
```

```{r}
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")
dn_lq_nc
```

```{r}
dn_lq_nc <- dn_lq_nc %>%
  mutate(dn_lq_nc, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_nc
```

336 observations, the variables are Address, city, zip, longitude, and latitude of both x and y and state

```{r}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_nc_mindist
```
```{r}
dn_lq_nc <- dn_lq_nc %>%
  mutate(dn_lq_nc, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_nc
```

```{r}
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_nc_mindist
```

```{r}
dn_lq_nc%>%
  ggplot(mapping = aes(x=address.x, y=distance)) + geom_point()
```


```{r}
dn_lq_nc_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```



### Exercise 10
```{r}
dn_tx <- dennys %>%
  filter(state == "TX")
nrow(dn_tx)
```

```{r}
lq_tx <- laquinta %>%
  filter(state == "TX")
nrow(lq_tx)
```

```{r}
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")
dn_lq_tx
```

47,400 observations, the variables are Address, city, zip, longitude, and latitude of both x and y (separately) and state

```{r}
dn_lq_tx <- dn_lq_tx %>%
  mutate(dn_lq_tx, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_tx
```

```{r}
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_tx_mindist
```


```{r}
dn_lq_tx_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```
As you can see here, it does appear as if there is a La Quinta very close to almost every Dennys, only a handful were farther away.


### Exercise 11


```{r}
dn_ca <- dennys %>%
  filter(state == "CA")
nrow(dn_ca)
```

```{r}
lq_ca <- laquinta %>%
  filter(state == "CA")
nrow(lq_ca)
```

```{r}
dn_lq_ca <- full_join(dn_ca, lq_ca, by = "state")
dn_lq_ca
```



```{r}
dn_lq_ca <- dn_lq_ca %>%
  mutate(dn_lq_ca, distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y, round = 3))
dn_lq_ca
```

```{r}
dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
dn_lq_ca_mindist
```



```{r}
dn_lq_ca_mindist%>%
  ggplot(mapping = aes(x=address.x, y=closest)) + geom_point()
```



### Exercise 12

The joke appears most applicable in California

